The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2016-08-23 script/en/chapters/movies.tex
CT Book ed 12: Major update -- many changes, corrections, and additions during last 4 months

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\newpage
\hypertarget{appendix-movies}{}%neu
\section{Movies and Fictional Literature with Relation to Cryptography}
\label{s:appendix-movies}
% {\bf movies and literature related to cryptography}(see appendix \ref{s:appendix-movies})
\index{movies}
\index{literature}

@@

% --------------------------------------------------------------------------
\subsection{For Grownups and Teenagers}

@@

\item[\textrm{[Poe1843]}] \index{Poe 1843}
    Edgar Allan Poe\index{Poe, Edgar Allan}, \\
    {\em The Gold Bug}, 1843. \\
    In this short story Poe tells as first-person narrator about his
    acquaintanceship with the curious Mr. Legrand. They detect a fabulous
    treasure via a gold bug and a vellum found at the coast of New England.\\
    The cipher consists of 203 cryptic symbols and it proves to be a
    general monoalphabetic substitution cipher (see
    chapter~\ref{monoalphabeticSubstitutionCiphers}).
    The story tells how they solve the riddle step by step using a combination
    of semantic and syntax analysis (frequency analysis of single letters in
    English texts).\\
    In this novel the code breaker Legrand says the famous statement:
    ``Yet it may be roundly asserted that human ingenuity cannot concoct a
    cipher which human ingenuity cannot resolve -- given the according
    dedication.''\\
    % Yet it may be roundly asserted that human ingenuity cannot concoct a
    % cipher which human ingenuity cannot resolve...

\item[\textrm{[Verne1885]}] \index{Verne 1885}
    Jules Verne\index{Verne, Jules}, \\
    {\em Mathias Sandorf}, 1885. \\
    This is one of the most famous novels of the French author Jules Verne
    (1828-1905), who was called ``Father of Science fiction''.\\
    In ``Mathias Sandorf'' he tells the story of the freedom fighter Earl
    Sandorf, who is betrayed to the police, but finally he can escape.\\
    The whistle-blowing worked, because his enemies captured and decrypted
    a secret message sent to him. For decryption they needed a special
    grille, which they stole from him. This turning grille was a quadratic
    piece of jig with 6x6 squares, of which 1/4 (nine) were holes
    (see the \hyperlink{turning-grille}{turning grille} in 
    chapter~\ref{introsamplesTranspositionCiphers}).\\

@@

\item[\textrm{[Seed1990]}] \index{Seed 1990}
    Directed by Paul Seed, \\
    {\em House of Cards}, 1990. \\
    In this movie Ruth tries to solve the secret, which made her daughter
    fall silent. Here two young people suffering from autism communicate
    via 5- and 6-digit primes (see 
    chapter~\ref{Label_Kapitel_Primes}).
    After more than 1 hour the movie contains the following encrypted
    two series of primes:
    \begin{center}
    $21,383; \;\;176,081; \;\;18,199; \;\;113,933; \;\;150,377; \;\;304,523; \;\;113,933$ \\
    $193,877; \;\;737,683; \;\;117,881; \;\;193,877$
    \end{center}
    % where autistic children communicate via primes
    \vskip +20 pt

\item[\textrm{[Robinson1992]}] \index{Robinson 1992}
    Directed by Phil Alden Robinson, \\
    {\em Sneakers}, Universal Pictures Film, 1992. \\
    In this movie the ``sneakers'', computer experts under their boss Martin
    Bishop, try to get back the deciphering box SETEC from the ``bad guys''.
    SETEC, invented by a genius mathematician before he was killed, allows to
    decrypt all codes from any nation. \\
    In the movie the code is not described in any way%
    \footnote{
       Leonard Adleman (the "A" within RSA) worked as mathematical consultant
       for ``Sneakers''. He describes the funny story about his contribution
       at his homepage 
       \url{http://www.usc.edu/dept/molecular-science/fm-sneakers.htm}.
       It is assumed that the cipher used everywhere is RSA.      
       According to that within the chip a fast, unknown factorization
       method\index{factorization} is implemented.
    }.\\

@@

\item[\textrm{[Brown1998]}] \index{Brown 1998}
    Dan Brown, \\
    {\em Digital Fortress}, E-Book, 1998. \\
    Dan Brown's first novel was published in 1998 as e-book, but it was
    largely unsuccessful then.\\
    The National Security Agency (NSA) uses a huge computer, which enables it
    to decrypt all messages (needless to say only of criminals and terrorists)
    within minutes even if they use the most modern encryption methods.\\
    An apostate employee invents an unbreakable code and his computer program
    Diabolus forces the super computer to do self destructing operations.
    The plot, where also the beautiful computer expert Susan Fletcher has a
    role, is rather predictable.\\
    The idea, that the NSA or another secret service is able to decrypt any
    code, is currently popular on several authors: In ``Digital Fortress'' the
    super computer has 3 million processors -- nevertheless from today's sight
    this is by no means sufficient to hack modern ciphers.\\

@@

\item[\textrm{[Suarez2010]}] \index{Suarez 2010}
    Daniel Suarez, \\
    {\em Freedom (TM)}, Penguin Books, 2010, 486 pages. \\
    ``The propulsive, shockingly plausible sequel to the bestseller Daemon'' (see above).
    Freedom (TM) (Daemon \#2) patches a number of holes the writer left in the first book.
    The prose is tighter, the descriptions more direct, the characters are fleshed out,
    especially Loki.
    Having laid the groundwork in ``Daemon'', Suarez uses this
    foundation in order to explore a new concept of social organization based on empowering
    information technology and the reasoning why and how the battle runs between the old
    potentates and the Daemon society, which also evolves further already during the story.
    Cryptography is a natural part of modern technology and modern warfare as described
    in this book.
    The new society emerging in "Freedom (TM)" is based on the darknet, an alternative to
    the internet using fast wireless meshes in order to increase the durability and
    availability of the network. Despite the story is shocking in some parts, it appears to
    be realistic and not far away from the parallel usage of modern technology integrated
    into our modern lives as a virtual world overlaying our real world.\\

@@

\item[\textrm{[Elsberg2012]}] \index{Elsberg 2012}
    Marc Elsberg, \\
    {\em Blackout -- Tomorrow its too late (original title: Blackout -- Morgen ist
    es zu sp\"at)},\\ Blanvalet, 2012, 800 pages. \\
    At a could day in winter all power supply networks in Europe break down.
    Agencies, energy suppliers and security companies are in the dark and unable to solve
    the problem.
    The Italian computer scientist Piero Manzano believes that this caused by terrorists
    using hackers: All customers use since some years smart meters, electricity meters
    controlled by software which was manipulated. Despite the integrated security and encryption
    components they have been hacked, and are out of order by wrong control sequences.
    The terrifying consequences at various locations are described realistically and excitingly.
    And in the same way the reactions of the human beings ...\\

@@

\vskip +20 pt
\paragraph*{Remark 1:}
Further samples of cryptology in fictional literature can be found on the
following German web page:
\parskip -4pt  % 1-BACK to DEFAULT -- Really necessary
\begin{center}
    \url{http://www.staff.uni-mainz.de/pommeren/Kryptologie99/Klassisch/1\_Monoalph/Literat.html}
\end{center}
\parskip \value{mycounterDefaultParskip} pt  % 2-BACK to DEFAULT -- Really necessary
For some older authors (e.g. Jules Verne, Karl May, Arthur Conan Doyle,
Edgar Allen Poe) there are links to the original and relevant text pieces.\\

\paragraph*{Remark 2:}
You can find title pages of some of these books on the web site of
Tobias Schr\"odel\index{Schroedel, Tobias}, who collects classic books about cryptography:
\parskip -4pt  % 1-BACK to DEFAULT -- Really necessary
\begin{center}
   \url{http://tobiasschroedel.com/crypto_books.php}
\end{center}
\parskip \value{mycounterDefaultParskip} pt  % 2-BACK to DEFAULT -- Really necessary
\vspace{1em}

\paragraph*{Remark 3:}
If you know of further books and movies, where cryptography has a major
role then we would be very glad if you could send us the exact title and
a short explanation about the movie/book's content. Thanks a lot.\\

@@

    Right at page 6 it summarizes for beginners in an old fashion style
    ``The Important Things First'' about paper\&pencil encryption
    (compare chapter~\ref{Kapitel_PaperandPencil}):
    \begin{itemize}
      \item[-] ``It must be possible to encrypt your messages at any place and
               at any location with the easiest measures and a small effort
               in a short time.
      \item[-] Your cipher must be easy to remember and easy to read for your
               partners. But strangers should not be able to decrypt them.\\
               Remember: Fastness before finesse, security before carelessness.
      \item[-] Your message must always be as short and precise as a telegram.
               Shortness outranks grammar and spelling. Get rid of all needless
               like salutations or punctuation marks. Preferably use only
               small or only capital letters.''
    \end{itemize}
    \vskip +20 pt   % da "\\" hier nicht geht!

@@

@@

CT1\index{CrypTool 1} contains an interactive educational tool for elementary
{\em number theory}, called ``NT''.\footnote{%
    NT can be called in CT1\index{CrypTool 1} via the menu path
    {\bf Indiv. Procedures \textbackslash{} Number Theory Interactive
    \textbackslash{} Learning Tool for Number Theory}.
}

@@

The content of this educational tool is mainly based on the books by J.\ Buchmann and H.\ Scheid
\cite{a:Buchmann2004,a:Scheid2003}.

@@

% -> Figur 1  WIE ERREICHT MAN, dass er von vorne zählt ? (nicht wichtig)
\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C1-3_EuclidsAlg-LinearCombinations.png}
\caption{Each common divisor of two integers also divides all its linear
combinations\vspace{1ex}} 
\label{NT_Fig_C1.3_EuclidsAlg-LinearCombinations}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C1-3_EuclidsAlg-Procedure}
\caption{Euklid's algorithm to determine gcd\vspace{1ex}} 
\label{NT_Fig_C1.3_EuclidsAlg-Procedure}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C3-1_PrimesDistribution}
\caption{Distribution of primes and its differences\vspace{1ex}} 
\label{NT_Fig_C3.1_PrimesDistribution}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C3-2_Fermat-Test}
\caption{Finding primes with the prime number test of Fermat\vspace{1ex}} 
\label{NT_Fig_C3.2_Fermat-Test}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C4-1_ReversibilityAdditiveCipher}
\caption{Reversibility of encryption mechanisms exemplified with additive
ciphers\vspace{1ex}} 
\label{NT_Fig_C4.1_ReversibilityAdditiveCipher}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C5-2_Fermat-factorization-How-far-1}
\caption{Fermat factorization using the third binomial formula\vspace{1ex}} 
\label{NT_Fig_C5.2_Fermat-factorization-How-far-1}
\end{center}
\end{figure}

>>>>

% $Id: 
% ..............................................................................
%             M O V I E S  +  N O V E L S
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\begin{bibunit}[babalpha] %% alpha: Chapter bibliography shows authors abbreviation

\newpage
\hypertarget{appendix-movies}{}%neu
\section{Movies and Fictional Literature with Relation to Cryptography}
\label{s:appendix-movies}
% {\bf movies and literature related to cryptography}(see appendix \ref{s:appendix-movies})
\index{movies}
\index{literature}

@@

% --------------------------------------------------------------------------
\subsection{For Grownups and Teenagers}
\label{s:Light-fiction-for-grownups}

@@

\item[\textrm{[Poe1843]}] \index{Poe 1843}
    Edgar Allan Poe\index{Poe, Edgar Allan}, \\
    {\em The Gold Bug}, 1843.\footnote{%
        See \url{https://en.wikipedia.org/wiki/The_Gold-Bug}.

	Part 1 of the series {\em RSA \& Co. at school: Modern cryptology,
        old mathematics, and subtle protocols} contains a didactical description
	for usage at school.
	Unfortunately this series is currently only available in German.
        See \cite{Witten1998}, pp 52 ff (``Das Gold des Gehenkten'').
	\index{RSA \& Co. at school}

	All material about a gold-bug lesson (double period) can be found at
	\url{http://www.informatik-im-kontext.de/} via
	``E-Mail (nur?) für Dich'' => ``Vertraulichkeit mit Verschlüsselungsverfahren''.

	Poe not only was a well-known writer, but also a talented cryptographer.
	His story is also told in the book {\em Code Breaking} \cite{Kippenhahn1997}.
    }\\
    In this short story Poe tells as first-person narrator about his
    acquaintanceship with the curious Mr. Legrand. They detect the fabulous
    treasure of captain Kidd via a gold bug and a vellum found at the coast
    of New England.\\
    The cipher consists of 203 cryptic symbols and it proves to be a
    general monoalphabetic substitution cipher (see
    chapter~\ref{monoalphabeticSubstitutionCiphers}).
    The story tells how they solve the riddle step by step using a combination
    of semantic and syntax analysis (frequency analysis of single letters in
    English texts).\footnote{
      Part of the learning material above is a Python program which can be used
      to decrypt the ciphertext with Python {\em and} with SageMath\index{SageMath}.
      See the code example in \ref{s:appendix-Code-for-light-fiction-books}.
    }\\
    In this novel the code breaker Legrand says the famous statement:
    ``Yet it may be roundly asserted that human ingenuity cannot concoct a
    cipher which human ingenuity cannot resolve -- given the according
    dedication.''\\
    % Yet it may be roundly asserted that human ingenuity cannot concoct a
    % cipher which human ingenuity cannot resolve...

\item[\textrm{[Verne1885]}] \index{Verne 1885}
    Jules Verne\index{Verne, Jules}, \\
    {\em Mathias Sandorf}, 1885. \\
    This is one of the most famous novels of the French author Jules Verne
    (1828-1905), who was called ``Father of Science fiction''.\\
    In ``Mathias Sandorf'' he tells the story of the freedom fighter Earl
    Sandorf, who is betrayed to the police, but finally he can escape.\\
    The whistle-blowing worked, because his enemies captured and decrypted
    a secret message sent to him. For decryption they needed a special
    grille, which they stole from him. This turning grille was a quadratic
    piece of jig with 6x6 squares, of which 1/4 (nine) were holes
    (see the \hyperlink{turning-grille-cipher}{turning grille} in 
    chapter~\ref{introsamplesTranspositionCiphers}).\\

@@

\item[\textrm{[Seed1990]}] \index{Seed 1990}
    Directed by Paul Seed, \\
    {\em House of Cards}, 1990. \\
    In this movie Ruth tries to solve the secret, which made her daughter
    fall silent. Here two young people suffering from autism communicate
    via 5- and 6-digit primes (see 
    chapter~\ref{Chapter_Primes}).
    After more than 1 hour the movie contains the following encrypted
    two series of primes:
    \begin{center}
    $21,383; \;\;176,081; \;\;18,199; \;\;113,933; \;\;150,377; \;\;304,523; \;\;113,933$ \\
    $193,877; \;\;737,683; \;\;117,881; \;\;193,877$
    \end{center}
    % where autistic children communicate via primes
    \vskip +20 pt

\item[\textrm{[Robinson1992]}] \index{Robinson 1992}
    Directed by Phil Alden Robinson, \\
    {\em Sneakers}, Universal Pictures Film, 1992. \\
    In this movie the ``sneakers'', computer experts under their boss Martin
    Bishop, try to get back the deciphering box SETEC from the ``bad guys''.
    SETEC, invented by a genius mathematician before he was killed, allows to
    decrypt all codes from any nation. \\
    In the movie the code is not described in any way%
    \footnote{
       Leonard Adleman (the ``A'' within RSA) worked as mathematical consultant
       for ``Sneakers''. He describes the funny story about his contribution
       at his homepage 
       \url{http://www.usc.edu/dept/molecular-science/fm-sneakers.htm}.
       It is assumed that the cipher used everywhere is RSA.      
       According to that within the chip a fast, unknown factorization
       method\index{factorization} is implemented.
    }.\\

@@

\item[\textrm{[Brown1998]}] \index{Brown 1998}
    Dan Brown, \\
    {\em Digital Fortress}, E-Book, 1998. \\
    Dan Brown's first novel was published in 1998 as e-book, but it was
    largely unsuccessful then.\\
    The National Security Agency (NSA) uses a huge computer, which enables it
    to decrypt all messages (needless to say only of criminals and terrorists)
    within minutes even if they use the most modern encryption methods.\\
    An apostate employee invents an unbreakable code and his computer program
    Diabolus forces the super computer to do self destructing operations.
    The plot, where also the beautiful computer expert Susan Fletcher has a
    role, is rather predictable.\\
    The idea, that the NSA or another secret service is able to decrypt any
    code, is currently a popular topic. % popular on several authors:
    In ``Digital Fortress'' the super computer has 3 million processors -- 
    nevertheless from today's view this is by no means sufficient to hack
    modern ciphers.\\

@@

\item[\textrm{[Suarez2010]}] \index{Suarez 2010}
    Daniel Suarez, \\
    {\em Freedom (TM)}, Penguin Books, 2010, 486 pages. \\
    ``The propulsive, shockingly plausible sequel to the bestseller Daemon'' (see above).
    Freedom (TM) (Daemon \#2) patches a number of holes the writer left in the first book.
    The prose is tighter, the descriptions more direct, the characters are fleshed out,
    especially Loki.
    Having laid the groundwork in ``Daemon'', Suarez uses this
    foundation in order to explore a new concept of social organization based on empowering
    information technology and the reasoning why and how the battle runs between the old
    potentates and the Daemon society, which also evolves further already during the story.
    Cryptography is a natural part of modern technology and modern warfare as described
    in this book.
    The new society emerging in ``Freedom (TM)'' is based on the darknet, an alternative to
    the internet using fast wireless meshes in order to increase the durability and
    availability of the network. Despite the story is shocking in some parts, it appears to
    be realistic and not far away from the parallel usage of modern technology integrated
    into our modern lives as a virtual world overlaying our real world.\\

\item[\textrm{[Olsberg2011]}] \index{Olsberg 2011}
    Karl Olsberg,\\
    {\em Rafael 2.0 (original title: Rafael 2.0)}, Thienemann Verlag, 2011, 240 pages.\\
    Michael and Rafael Ogilvy are talented twins who get along very well.
    Before the terminally ill Rafael dies, his father developed a virtual
    computer effigy of him, an artificial intelligence (AI). This is a good
    kept secret until Michael one day finds out what his father is hiding before
    him. However, his first horror soon turns into joy.
    So he still has something that reminds him of his brother. \\
    But this computer system is also interesting for the military. One day Michael's
    father is kidnapped and the company and thus also the computer program Rafael 2.0
    fall into the wrong hands.
    Michael is banished by his uncle in a boarding school, from which he can flee.
    Henceforth, Michael and his friends try their best to find his father, of whom
    they assume that he was abducted by a competing company. From here the
    story gets really exciting ...
    Michael learns that there is another artificial intelligence, Metraton,
    which is not so well-disposed to the people.
    Nothing is too much engrossed, young teenagers are the target audience.
    Nevertheless, depth and substance are created when for instance the machinations
    in acquisitions are discussed. \\
    From a crypto perspective,the section about factoring is thrilling: With a
    variant Michael can detect whether the computer is cheating ...\\

@@

\item[\textrm{[Eschbach2011]}] \index{Eschbach 2011}
    Andreas Eschbach, \\
    {\em Master of the universe -- master of all staff (original title: Herr aller Dinge)},
    Lübbe, 2011. \\
    This novel would have deserved a much broader audience: The idea in it
    of the ``most terrific of all crimes'', which is the origin of the whole
    story, is new and almost revolutionary, but also infinitely sad.
    Along the failing partnership of Hiroshi (inventor genius) and Charlotte
    important topics like justice, human wealth and power are dealt with.\\
    From a crypto perspective, Hiroshi uses distributed calculations and developed
    an encryption and backup system which misleads the government which buged him.\\

\item[\textrm{[Elsberg2012]}] \index{Elsberg 2012}
    Marc Elsberg, \\
    {\em Blackout -- Tomorrow its too Late (original title: Blackout -- Morgen ist
    es zu sp\"at)},\\ Blanvalet, 2012, 800 pages. \\
    At a could day in winter all power supply networks in Europe break down.
    Agencies, energy suppliers and security companies are in the dark and unable to solve
    the problem.
    The Italian computer scientist Piero Manzano believes that this caused by terrorists
    using hackers: All customers use since some years smart meters, electricity meters
    controlled by software which was manipulated. Despite the integrated security and encryption
    components they have been hacked, and are out of order by wrong control sequences.
    The terrifying consequences at various locations are described realistically and excitingly.
    And in the same way the reactions of the human beings ...\\

\item[\textrm{[Olsberg2013]}] \index{Olsberg 2013}
    Karl Olsberg,\\
    {\em The Eigths Revelation (original title: Die achte Offenbarung)},\\
    Aufbau Taschenbuch, 2013, 460 pages.\\
    Can a message from the past change our future?
    An ancient, encrypted manuscript fell into the hands the historian Paul Brenner.
    The more he decodes the text, the more puzzling is the content: Because the book
    tells with remarkable precision events years ahead of the time of its presumed creation.
    While highly dangerous genetic material disappears from a US laboratory, someone
    tries to prevent at any price, that Paul deciphers the last, the eighth revelation.
    A gripping thriller about a shockingly realistic apocalypse with many human aspects ...\\
    As a reader, you can participate in the deciphering of the manuscript.\\
    The experiments of Paul to make the right persons aware of his discovery and to
    correct it later, are described very exciting -- even chief editors have a dilemma
    with conspiracy.\\
    The cipher on the last book page is offered as a challenge in the crypto
    competition MTC3:
    \url{https://www.mysterytwisterc3.org/en/challenges/level-i/the-last-note}\\
    % https://www.mysterytwisterc3.org/images/challenges/mtc3-esslinger-18-note-en.pdf

\item[\textrm{[Takano2014]}] \index{Takano 2014}
    Kazuaki Takano, \\
    {\em Genocide of One}, 2014. (Orginal in Japanisch: ``Jenosaido'', 2011;
    as paperback in Englisch again under the title ``Extinction'', 2016) \\
    The cover text says: He is a new kind of human. He may mean the end for the rest of us...
    One bright morning in Washington D.C., the US President learns of a terrifying new
    threat to national security. Soon afterwards, American mercenary Jonathan Yeager is
    asked to lead a team into the Congo to eliminate a mysterious enemy -- a job which
    will help him pay for treatment for his dying son. But when they reach Africa, the
    threat turns out to be a three-year-old child named Akili: the next step in human
    evolution. The soldiers are under orders to kill the boy before his full potential
    can be realized. Yet Akili's advanced knowledge might be the only hope Yeager has
    to save his son's life... With time running out to choose a side, Yeager must
    decide whether to follow his orders or to save a creature who may not be as
    harmless or innocent as he appears. Because Akili is already the smartest being
    on the planet, with the power to either save humanity -- or destroy it.\\
    This is a very exciting book. After having overcome the first 100-200 pages
    you'll be awarded with surprising insights. According to the recensions its
    very well researched, but not for superficial readers.\\
    From a crypto perspective, RSA and OTP are directly drivers of the story and
    are explained correctly. Breaking RSA by factorization is so important that
    the CIA wouldn't accept that this knowledge isn't in their ownership ...\\

\item[\textrm{[Elsberg2014]}] \index{Elsberg 2014}
    Marc Elsberg,\\
    {\em ZERO -- They Know what you are Doing (original title: ZERO -- Sie wissen, was du tust)},\\
    Blanvalet Verlag, 2014, 480 pages.\\
    London. In a pursuit a boy is shot. His death takes the journalist Cynthia Bonsant
    to the acclaimed internet platform Freemee. Freemee collects and analyzes data,
    and thus promises its millions of users -- rightly -- a better life and more
    success. There is only one who warns about Freemee and about the power that the
    online newcomer could give just a few: ZERO, the most searched online activist
    in the world. As Cynthia begins precisely to research, she's becoming the quarry.
    And in a world of cameras, headsets and smartphones there is no escape ... \\
    Highly topical and menacing: the transparent person under control.
    The novel takes place in the near future (near fiction) and contains many
    contemporary references such as PRISM, predictive analytics, gamification.
    By the way, references to well-known science fiction media like ``The Running Man'',
    ``Monkey Wrench Gang'', ``V as Vendetta'' (V wears a Guy Fawkes mask, now the
    hallmark of Anonymous), ``Network'' and ``Body Snatchers'' are processed. \\
    Technologically / cryptologically the protagonists move on the highest
    level, which is not further explained: Alice Kinkaid communicates with a
    Raspberry Pi. Cynthia's daughter Vi uses mesh networks. ~See\\
    \url{https://de.wikipedia.org/wiki/Zero_%E2%80%93_Sie_wissen,_was_du_tust},\\
    \url{http://www.zero-das-buch.de/actiontrailer.php}\\

\item[\textrm{[Lagercrantz2015]}] \index{Lagercrantz 2015}
    David Lagercrantz, \\
    {\em The Girl in the Spider's Web}, Quercus, 2015. \\
    This is the 4th novel in the Millennium series, and the first not written by
    Stieg Larsson. While Mikael Blomkvist's print medium is struggling to
    survive the reader gets more and more insight in the inner structures and the
    combinations of publishers, secret services, public agencies, organized
    crime, and industrial espionage.
    Here, no care is taken for single humans, and normal humans would have no
    chance aginst this mix of interests. However, the special skills of
    Lisbeth Salander makes a difference, and so the NSA is informed, that
    parts of it are led and misused  by organized crime. The characters of the
    Millenium trilogy have been developed further in a credible way. Very exiting.\\
    From a crypto perspective, Lisbeth and the autist August deal with elliptic
    curves to crack RSA.\\
    See \url{https://en.wikipedia.org/wiki/The_Girl_in_the_Spider%27s_Web}.\\

@@

\vskip +20 pt
\paragraph*{Remark 1:}
A long list of (partly commented) samples of cryptology in fictional literature can
be found on the following German web page:
\parskip -4pt  % 1-BACK to DEFAULT -- Really necessary
\begin{center}
    \url{http://www.staff.uni-mainz.de/pommeren/Kryptologie/Klassisch/0_Unterhaltung/}
\end{center}
\parskip \value{mycounterDefaultParskip} pt  % 2-BACK to DEFAULT -- Really necessary
For some older authors (e.g. Jules Verne, Karl May, Arthur Conan Doyle,
Edgar Allen Poe) there are links to the original and relevant text pieces.\\\\

\paragraph*{Remark 2:}
You can find title pages of some of these books on the web site of
Tobias Schr\"odel\index{Schroedel, Tobias}, who collects classic books about cryptography:
\parskip -4pt  % 1-BACK to DEFAULT -- Really necessary
\begin{center}
   \url{http://tobiasschroedel.com/crypto_books.php}
\end{center}
\parskip \value{mycounterDefaultParskip} pt  % 2-BACK to DEFAULT -- Really necessary
\vspace{2em}

\paragraph*{Remark 3:}
If you know of further books and movies, where cryptography has a major
role then we would be very glad if you could send us the exact title and
a short explanation about the movie/book's content. We will insinuate
your possible enthusiasm for a title. Thanks a lot.\\

@@

    Right at page 6 it summarizes for beginners in an old fashion style
    ``The Important Things First'' about paper\&pencil encryption
    (compare chapter~\ref{Chapter_PaperandPencil}):
    \begin{itemize}
      \item[-] ``It must be possible to encrypt your messages at any place and
               at any location with the easiest measures and a small effort
               in a short time.
      \item[-] Your cipher must be easy to remember and easy to read for your
               partners. But strangers should not be able to decrypt them.\\
               Remember: Fastness before finesse, security before carelessness.
      \item[-] Your message must always be as short and precise as a telegram.
               Shortness outranks grammar and spelling. Get rid of all needless
               like salutations or punctuation marks. Preferably use only
               small or only capital letters.''
    \end{itemize}
    \vskip +20 pt   % da "\\" hier nicht geht!

@@

% --------------------------------------------------------------------------
\newpage
\subsection{Code for the light fiction books}
     \label{s:appendix-Code-for-light-fiction-books}

\noindent Chapter \ref{s:Light-fiction-for-grownups} lists as first book
``The Gold Bug'' by E.A. Poe.\index{Poe 1843}

Using the following Python code\footnote{%
  See \cite{Witten1998} and the file \verb|2_MonoKidd.zip| at
  \url{http://bscw.schule.de/pub/bscw.cgi/159132}.
}
you can decrypt the ciphertext of Captain Kidd (see the original text of
``The Gold Bug'' in \url{http://pinkmonkey.com/dl/library1/gold.pdf},
page 21) with Python or with SageMath\index{SageMath}.

The code already contains the ascii characters of the ciphertext and the
correlated alphabets for the plaintext and the ciphertext of this
monoalphabetic cipher.

The easiest way to perform the decryption is using the SageMathCell server
(\url{http://sagecell.sagemath.org/}) in a browser:
There you can switch between the programming languages Sage and Python.
The code can be executed by insering it with ``copy and paste'' and
then pressing ``Evaluate''.\\

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize]

# Hier geht's los ... Start here
import string
PA = 'ETHSONAIRDGLBVPFYMUC'
print 'Plaintext alphabet  PA: ', PA, '   Length of PA', len(PA)
CA = "8;4)+*56(!302'.1:9?-"
print 'Ciphertext alphabet CA: ', CA, '   Length of CA', len(CA)
codetableC2P = string.maketrans(CA,PA)
C = '''53++!305))6*;4826)4+.)4+);806*;48!8'60))85;1+(;:+*8!83(88)5*!;46
(;88*96*?;8)*+(;485);5*!2:*+(;4956*2(5*-4)8'8*;4069285);)6!8)4++;1(+9;4
8081;8:8+1;48!85;4)485!528806*81(+9;48;(88;4(+?34;48)4+;161;:188;+?;'''
P = string.translate(C, codetableC2P);
print 'Kidd decrypted:', P
# ... und hier sind wir schon fertig ... and here we are done!

\end{Verbatim}
\caption{Decryption of the Gold-Bug ciphertext from the novel of E.A. Poe (with Python)}
\label{Lit_Python-sample_Gold-bug}
\end{sagecode}

\paragraph*{Remark 1:}
When printing the ciphertext of Poe, Pinkmonkey ``cheated'' similarly like
the code author by using only ascii characters.

In the archive of an original publication (e.g. at
\url{https://archive.org/details/goldbug00poegoog} at page 95) you can see,
that Poe used characters which were common in the letterpress printing 
(and most of them are also part of the ascii set). It is very unlikely
that an untaught pirate would use just such characters for his ciphertext ...

\paragraph*{Remark 2:}
The code uses the Python functions ``maketrans'' und ``translate'' from the
String package (see e.g. the Python 2.7 documentation, section ``functions'').

So both alphabets (for the plaintext and the ciphertext) are inserted as a simple
string, and ``maketrans'' creates a mapping table. The actual encryption
is done by ``translate''. For the decryption you just have to switch the
arguments of ``maketrans'' for the two alphabets.
The otherwise necessary transformations between characters and their ascii numbers
(using ``str'' and ``ord'') can be avoided. This is ideal for
monoalphabetic ciphers -- especially for lessons at the junior high school.

\paragraph*{Remark 3:}
The following screenshot shows the execution of the Python code on the SageMathCell
server (\url{http://sagecell.sagemath.org/}).

SageMathCell is the browser interface for Sage. Its functions are delivered for free
on the Sage cell server (Sage is a comprehensive open-source math software system).

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.53]{figures/Using-SageMathCell-with-Python-for-Poes-GoldBug}
\caption{Usage of SageMathCell to decipher Poe's Gold Bug (with Python)} 
\label{MOV_Using-SageMathCell-with-Python-for-Poes-GoldBug}
\end{center}
\end{figure}

\noindent As result we get:

\begin{Verbatim}%

Plaintext alphabet  PA:  ETHSONAIRDGLBVPFYMUC    Length of PA 20
Ciphertext alphabet CA:  8;4)+*56(!302'.1:9?-    Length of CA 20
Kidd decrypted: AGOODGLASSINTHEBISHOPSHOSTELINTHEDEVILSSEATFORTYONEDEGREES
                ANDTHIRTEENMINUTESNORTHEASTANDBYNORTHMAINBRANCHSEVENTHLIMB
                EASTSIDESHOOTFROMTHELEFTEYEOFTHEDEATHSHEADABEELINEFROMTHE
                TREETHROUGHTHESHOTFIFTYFEETOUT

\end{Verbatim}

\noindent It's evident how less code is needed with Python or Sage for such tasks.
In the above sample there were 2 unnecessary comment lines, 3 lines for input,
3 lines of real code, and 3 lines for the output; really necessary were only
7 lines of code.

@@

CT1\index{CT1} contains an interactive educational tool for elementary
{\em number theory}, called ``NT''.\footnote{%
    NT can be called in CT1\index{CT1} via the menu path
    {\bf Indiv. Procedures \textbackslash{} Number Theory Interactive
    \textbackslash{} Learning Tool for Number Theory}.
}

@@

The content of this educational tool is mainly based on the books by
J.\ Buchmann and H.\ Scheid \cite{Buchmann2016, Scheid2006}.

@@

% -> Figur 1  WIE ERREICHT MAN, dass er von vorne zählt ? (nicht wichtig)
\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C1-3_EuclidsAlg-LinearCombinations.png}
\caption{Each common divisor of two integers also divides all its linear combinations} 
\label{NT_Fig_C1.3_EuclidsAlg-LinearCombinations}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C1-3_EuclidsAlg-Procedure}
\caption{Euklid's algorithm to determine gcd} 
\label{NT_Fig_C1.3_EuclidsAlg-Procedure}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C3-1_PrimesDistribution}
\caption{Distribution of primes and its differences} 
\label{NT_Fig_C3.1_PrimesDistribution}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C3-2_Fermat-Test}
\caption{Finding primes with the prime number test of Fermat} 
\label{NT_Fig_C3.2_Fermat-Test}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C4-1_ReversibilityAdditiveCipher}
\caption{Reversibility of encryption mechanisms exemplified with additive
ciphers} 
\label{NT_Fig_C4.1_ReversibilityAdditiveCipher}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C5-2_Fermat-factorization-How-far-1}
\caption{Fermat factorization using the third binomial formula} 
\label{NT_Fig_C5.2_Fermat-factorization-How-far-1}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{figures/NT_Fig_C5-2_Fermat-factorization-How-far-2}
\caption{Fermat factorization: detecting squares} 
\label{NT_Fig_C5.2_Fermat-factorization-How-far-2}
\end{center}
\end{figure}
